A re-examination of the role of friction in the original Social Force Model

### 2019-05-22 paper.tex
quite las figuras de velocity profile (salvo la scaled velocity profile)

The desired velocity for each pedestrian $i$ was 
$\vec{v}_d^{~(i)}=1$~m/s~$\hat{e}_d^{~(i)}$, where the target $\hat{e}_d^{~(i)}$ 
was set as $\hat{e}_d^{~(i)}=(L,y_i)\left \| (L,y_i) \right \|^{-1}$, being $L$ 
the \textit{x}-location of the end at the corridor and $y_i$ the 
\textit{y}-location corresponding to the \textit{ith} pedestrian 
(\textcolor{red}{see Fig.~\ref{pasillo}}). This allowed the pedestrians to 
move from left to right in an unidirectional flow. Pedestrians that surpassed 
$x=L$ were re-injected at $x=0$, preserving their current velocity and 
\textit{y}-location (\textit{i.e.} periodic boundary conditions). This mechanism 
was carried out in order to keep
the crowd size unchanged.\\

@@

Fig.~\ref{speed-profile-w22} shows the velocity profile (velocity vs. \textit{y}-location) of the pedestrians across the corridor (see caption for details). We can see that low-density situations lead to a cruising velocity profile $v=v_d$. This is valid for every location in the corridor (not only the center as was previously noticed in Section \ref{fundamental_diagram}). For higher densities, the velocity profile turns into a parabola-like function. This shape resembles the usual velocity profile for laminar flow in a viscous fluid, where the velocity increases toward the center of a tube. In our case, pedestrians near the walls are the ones with the lower velocity. The velocity increases when departing from the wall until it reaches the maximum at the center of the corridor. This behavior suggests that the wall friction on the pedestrians, is playing a relevant role on the velocity distribution.\\

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./v_y_width22_k24.eps}
\caption{\label{speed-profile-w22} Mean velocity profile (velocity vs \textit{y}-position) for different densities (see the inserted legend). The simulated corridor was 28~m length. Pedestrians walk from left to right with periodic boundary condition in the \textit{x}-direction. Initially, pedestrians were randomly distributed, the corridor width was $w = 22~$m for all the cases. The bin size was 1~m.}
\end{figure}

Fig.~\ref{speed-profile-multi_width} shows the velocity profile for different widths. The horizontal axis of the plot corresponds to the $y$-location normalized by the width of the corridor. The density chosen was $\rho = 6$ since we wanted to study a situation in which pedestrians slow down. Recall that when $\rho<5$, collisions between pedestrians are not relevant (within the SFM). We can see that the lowest velocities occur in the regions near the walls. Additionally, there is a clear relation between $v_{max}$ and the corridor width. That is, the wider the corridor, the higher the maximum reached velocity.\\

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.7\columnwidth]
{./v_y_multi_width.eps}
\caption{\label{speed-profile-multi_width} Velocity profile (velocity vs $y$-position) for different corridors width (see legend for the corresponding widths). The simulated corridor was 28~m length. Pedestrians walk from left to right with periodic boundary condition in the $x$-direction. Initially, pedestrians were randomly distributed, the density was $\rho = 6$ in all the cases (high density regime). The bin size was 1~m except for $w=4$~m since the bin was 0.5~m.}
\end{figure}

Fig.~\ref{speed-profile-width-normaliz} exhibits the scaled velocity profile. The horizontal axis is normalized by the corresponding corridor width (just like in Fig.~\ref{speed-profile-multi_width}). Now, the vertical axis is normalized by the maximum velocity ($v_{max}$) corresponding to each data set. Filled markers correspond density $\rho=9$, while empty markers correspond to $\rho=6$. Notice that all the data follow the same  pattern, suggesting that the velocity profile exhibits a somewhat fundamental behavior, regardless the scale of the corridor (and the density). Hence, the velocity growth rate from the wall towards the center of the corridor, is the same in spite of the size of the corridor width and the density. \\

@@

@@

We may hypothesize that the friction force is somehow the key factor in the
flow reduction, as reported in the experimental fundamental diagram. This hypothesis further inspired us to analyze the role of the friction coefficient in a simple model for the corridor (see \ref{appendix_2}).

>>>>

The desired velocity for each pedestrian $i$ was 
$\vec{v}_d^{~(i)}=1$~m/s~$\hat{e}_d^{~(i)}$, where the target $\hat{e}_d^{~(i)}$ 
was set as $\hat{e}_d^{~(i)}=(L,y_i)\left \| (L,y_i) \right \|^{-1}$, being $L$ 
the \textit{x}-location of the end at the corridor and $y_i$ the 
\textit{y}-location corresponding to the \textit{ith} pedestrian 
(\textcolor{red}{see Fig.~\ref{corridor}}). This allowed the pedestrians to 
move from left to right in an unidirectional flow. Pedestrians that surpassed 
$x=L$ were re-injected at $x=0$, preserving their current velocity and 
\textit{y}-location (\textit{i.e.} periodic boundary conditions). This mechanism 
was carried out in order to keep
the crowd size unchanged.\\

@@

We first noticed that low-density situations ($\rho<5$) lead to a cruising velocity profile $v=v_d$. This is valid for every location in the corridor (not only the center as was previously noticed in Section \ref{fundamental_diagram}). For higher densities($\rho>5$), the velocity profile turns into a parabola-like function. This shape resembles the usual velocity profile for laminar flow in a viscous fluid, where the velocity increases toward the center of a tube. In our case, pedestrians near the walls are the ones with the lower velocity. The velocity increases when departing from the wall until it reaches the maximum at the center of the corridor. This behavior suggests that the wall friction on the pedestrians, is playing a relevant role on the velocity distribution.\\

Fig.~\ref{speed-profile-width-normaliz} exhibits the scaled velocity profile. The horizontal axis is normalized by the corresponding corridor width. The vertical axis is normalized by the maximum velocity ($v_{max}$) corresponding to each data set. Filled markers correspond density $\rho=9$, while empty markers correspond to $\rho=6$. Notice that all the data follow the same  pattern, suggesting that the velocity profile exhibits a somewhat fundamental behavior, regardless the scale of the corridor (and the density). Hence, the velocity growth rate from the wall towards the center of the corridor, is the same in spite of the size of the corridor width and the density. \\

@@

We want to mention that there is a clear relation between $v_{max}$ and the corridor width. That is, the wider the corridor, the higher the maximum reached velocity. (Fig.~\ref{speed-profile-width-normaliz} does not exhibit this behavior because the velocity is normalized by $v_{max}$).\\

@@
